---
title: "Assignment_3"
author: "Johnny Rodriguez"
date: "2022-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major, provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r}

# Load the libraries
library(RCurl)
library(tidyverse)
library(stringr)

# Gets the CSV from Github
majorscsv <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

# Reads the majors CSV into a dataframe.
majorsdf <- read.csv(text=majorscsv)

# To perform the matching in the Majors dataframe column (and avoid a coercion warning), the df is subset and vectorized.
majorsvec <- c(majorsdf[, "Major"])

# The strings that match 'data' and 'statistics' are extracted and assigned back to a clean dataframe. NA values are omitted.
majorsdf <- na.omit(data.frame(
  str_extract(majorsvec, regex((".*(data|statistic).*"), ignore_case = TRUE))))

# The final result shows 3 matches
majorsdf
```

#### Write code that transforms the data below

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}

# Creates the fruit vector
fruit <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

# Adds quotes to the elements in the fruit vector in preparation for the string collapse.
fruit <- paste(shQuote(fruit))

#Collapses the fruit vector into a single string
fruit <- str_c(fruit, collapse = ",")

# Concatenates the 'c' syntax to the fruit string
mystring <- paste('c(', fruit, ')')

#The new string is in a format that will recreate the original fruit vector.
mystring


```

#### Describe, in words, what these expressions will match:

This creates a single character capture group (.) followed by two back references itself. ie, repeats 2 more times.  Ex www.

(.)\1\1

This creates two, single character captures groups (.)(.) followed by a back reference to Group 2 \\2 followed by a back referenece to Group 1 \\1.  Ex.  abba
"(.)(.)\\2\\1"

This creates a single 2 character capture group (..) followed by a back reference to itself \1; ie, repeats itself once.  Ex. abab
(..)\1


This creates a single character capture group (.) followed by any character . followed by the back reference to the initial character \\1 followed by any character . followed by the back reference to the initial character. Ex a2axa
"(.).\\1.\\1"

This creates three single character capture groups (.)(.)(.) followed by any string of characters .* followed by the back reference to group 3 \\3 followed by a back reference to group 2 \\2 followed by a back reference to group 1.  Ex abc11111111cba
"(.)(.)(.).*\\3\\2\\1"


#### Construct regular expressions to match words that:
1. Start and end with the same character.
Creates a one character capture group that is asserted at the beginning of the string ^(.) followed by second capture group asserted.  This second capture group matches on any preceding character from the end with a back reference to the initial character \\1 asserted as the ending character $.

"^(.)(.*\\1$)"


2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Creates a 2-alpha character capture group ([A-Za-z][A-Za-z]) followed by any number of characters .* followed by a back reference to the initial 2 character capture group \\1

([A-Za-z][A-Za-z]).*\\1


3. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
Creates a single character capture group ([a-z]) followed by any number of characters .* followed by the back reference to the capture group \\1 followed by any number of character .* followed by a back reference to the capture group \\1.

([a-z]).*\\1.*\\1

